---
title: "Exam 2"
author: "Charles Etuk"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(lubridate)
library(dplyr)
library(tidyverse)
library(mdsr)
library(ggplot2)
library(plotly)
```
Download the data set found here: https://www.kaggle.com/datasets/arnabchaki/datascience-salaries-2023 Using this data set answer the questions below. Show all work!
```{r}
ds_salaries <- read.csv("/Users/charlesetuk/Downloads/ds_salaries.csv")
```
(a) (10 points) Find the average salary for all data science jobs by company size.
```{r}
ds_salaries %>% group_by(company_size) %>% summarise(avg_salary = mean(salary_in_usd))
```
(b) (10 points) Make this plot:
```{r}
ds_salariesUS <- ds_salaries %>% mutate(US = company_location == "US") %>% group_by(work_year, US) %>% summarise(mean_salary = mean(salary_in_usd))
ggplot(aes(x = work_year, y = mean_salary), data = ds_salariesUS) + geom_line(aes(color = US)) + geom_point(aes(color = US))
```


(c) (10 points) Using only companies in located in the US, find the mean
salary in USD by experience level and year. The table will look like
this:
```{r}
USsalaries <- ds_salaries %>% filter(company_location == "US") %>% group_by(experience_level, work_year) %>% summarise(mean_salary = mean(salary_in_usd)) 
USsalaries %>% pivot_wider(names_from = work_year, values_from = mean_salary)
```
(d) (20 points) Write a function that takes the company location as input
and returns the country and the salary in US dollars of the highest
paying full-time entry level position. Use the map family of functions
to run this function across all the company locations and create a data
set with two columns: company locations and salary. Remove rows
where the country had no entry level jobs in the data set and order
this data set from largest salary to lowest salary. Which locations
had the highest and lowest full time entry level positions?

```{r}
highestsal <- function(country) {
  data <- ds_salaries %>% 
    filter(experience_level == "EN" & employment_type == "FT") %>% filter(company_location == country)
  if (nrow(data) == 0) {
    return(NA)
  } else {
    topsalary <- data %>% 
      arrange(-salary_in_usd) %>% 
      head(1) %>% 
      pull(salary_in_usd)
    return(data.frame(country, topsalary))
  }
}

countrynames1 <- ds_salaries %>% filter(experience_level == "EN" & employment_type == "FT") 
countrynames <- unique(countrynames1$company_location)
results <- map_dfr(countrynames, highestsal)
results %>% arrange(-topsalary)
```
India (IN) had the highest top salary for full-time entry level positions and North Macedonia (MK) had the lowest top salary.


(e) (10 points) Find the mean salary in US dollars for all positions with the job title ”Data Engineer”. Using the bootstrap, estimate the standard error of this sample mean. Use at least 1000 bootstrap samples.
```{r}
dsengineers <- ds_salaries %>% filter(job_title == "Data Engineer")

mean(replicate(1000, sd(sample(dsengineers$salary_in_usd, replace=TRUE))/sqrt(length(dsengineers$salary_in_usd))))
```

2. Download the following two data sets:
CPI: https://www.kaggle.com/datasets/mikoajfish99/us-recession-and-financialindicators?select=Consumer+Price+Index+Total+All+Items+for+the+United+States.csv
Ten Year Real Interest Rate: https://www.kaggle.com/datasets/mikoajfish99/usrecession-and-financial-indicators?select=10-Year+Real+Interest+Rate.csv
```{r}
TenYrIR <- read.csv("/Users/charlesetuk/Downloads/10-Year_Real_Interest_Rate.csv")
CPI <- read.csv("/Users/charlesetuk/Downloads/Consumer_Price_Index_Total_All_Items_for_the_United_States.csv")

```
(a) (10 points) Plot the 10 year interest rate over time in an interactive
plot.
```{r}
IROverTime <- ggplot(aes(x = DATE, y = REAINTRATREARAT10Y), data = TenYrIR) + geom_point() + geom_line()
ggplotly(IROverTime)
```


(b) (10 points) Create side by side box plots of the consumer price index
grouped by month.
```{r}
CPIMonth <- CPI %>% mutate(month = month(DATE)) 
ggplot(aes(y = CPALTT01USM657N), data = CPIMonth) + geom_boxplot() + facet_wrap(~month)
```


(c) (20 points) Join the CPI data set and the IR data set on date. Create
a variable called ”decade” (i.e. 1980s, 1990s, 2000s, etc). Plot the
interest rate on the x-axis, the CPI on the y-axis, and use the decade
variable that you created as the color of each point. The plot should
look like this:
```{r}
TempPlot <- TenYrIR %>% left_join(CPI, by = c("DATE" = "DATE")) 
TempPlot$DATE <- as.Date(TempPlot$DATE) 
TempPlot2 <- TempPlot %>% mutate(decade = floor((year(TempPlot$DATE))/10) * 10)
TempPlot2$decade <- as.factor(TempPlot2$decade)
ggplot(aes(x = REAINTRATREARAT10Y, y = CPALTT01USM657N), data = TempPlot2) + geom_point(aes(color = decade))
```